---
title: "reproducible document"
author: "Melodie Schneider"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_ins = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```
## About me

- Mélodie Jil Schneider
- University of Copenhagen

**I** do *basic research*

## Simple code

```{r}
3 * 3
```
## Showing the data

```{r}
post_meal_data
```
```{r}
glimpse(post_meal_data)
```


```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Distribution of groups"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete variables

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## Putting tow plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glycemi
#| fig-cap: "BMI and glycemia side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

## Plotting two continous variables

```{r plotting two continous variables}
#| fig-cap: "Scatterplot of BMI and area under the curve of glycemia"
#| label: fig-scatterplot-bmi-aucglucose
#| echo: false
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```
## Pushing changes to remote

## Pulling changes locally

## Selecting data

```{r Select chunk}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming specific column

```{r Renaming specific column}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r Select and rename tasks}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5 = "P.Glucose..5.OGTT",
    Insulin.minus5 = "Insulin..5.OGTT.X"
  )

# Task 3.
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)

basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)
```

## filtering data by row

```{r filtering data by row}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")
```
```{r}
TRUE & TRUE
```
```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```
## Modify or add column
```{r Modify or add column}
tidier_post_meal_data |>
  mutate(
    insulin_glucose_ratio = auc_ins / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```

## Practice with filter and mutate

```{r Practice with filter and mutate}
# Task 1.
post_meal_data |>
  filter(BMI >= 20 & BMI <= 40 & Group == "FDR")

# Task 2.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```
